---
title: "Cross validation"
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(modelr)
library(mgcv)
library(rvest)
library(p8105.datasets)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.color = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_color_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d

set.seed(1)
```


## CV by hand

```{r}
data("lidar")

lidar_df = 
  lidar |> 
  as_tibble() |>
  mutate(id = row_number())

lidar_df |> 
  ggplot(aes(x = range, y = logratio)) + 
  geom_point()
```

### Split the data into training and test sets

```{r}
train_df = sample_frac(lidar_df, size = .8)
test_df = anti_join(lidar_df, train_df, by = "id")

ggplot(train_df, aes(x = range, y = logratio)) + 
  geom_point() + 
  geom_point(data = test_df, color = "red")
```

Fit three models

```{r}
linear_mod = lm(logratio ~ range, data = train_df)
smooth_mod = mgcv::gam(logratio ~ s(range), data = train_df)
wiggly_mod = mgcv::gam(logratio ~ s(range, k = 30), sp = 10e-6, data = train_df)
```

### Plot the two gam fits

```{r}
train_df |> 
  add_predictions(smooth_mod) |> 
  ggplot(aes(x = range, y = logratio)) + 
  geom_point() + 
  geom_line(aes(y = pred), color = "red")
```

```{r}
train_df |> 
  add_predictions(wiggly_mod) |> 
  ggplot(aes(x = range, y = logratio)) + 
  geom_point() + 
  geom_line(aes(y = pred), color = "red")
```

```{r}
train_df |> 
  gather_predictions(linear_mod, smooth_mod, wiggly_mod) |> 
  mutate(model = fct_inorder(model)) |> 
  ggplot(aes(x = range, y = logratio)) + 
  geom_point() + 
  geom_line(aes(y = pred), color = "red") + 
  facet_wrap(~model)
```

Look at prediction accuracy
## Compute root mean squared errors(RMSEs) for each model
```{r}
rmse(linear_mod, test_df)
```

```{r}
rmse(smooth_mod, test_df)
```

```{r}
rmse(wiggly_mod, test_df)
```


## CV using modelr

```{r}
cv_df = 
  crossv_mc(lidar_df, 100) 
```

What is happening here ... 

```{r}
cv_df |> pull(train) |> nth(1) |> as_tibble()
```

```{r}
cv_df |> pull(test) |> nth(1) |> as_tibble()
```

```{r}
cv_df =
  cv_df |> 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble))
```

I now have many training and testing datasets, and I’d like to fit my candidate models above and assess prediction accuracy as I did for the single training / testing split. To do this, I’ll fit models and obtain RMSEs using mutate + map & map2.

```{r}
cv_df = 
  cv_df |> 
  mutate(
    linear_mod  = map(train, \(df) lm(logratio ~ range, data = df)),
    smooth_mod  = map(train, \(df) gam(logratio ~ s(range), data = df)),
    wiggly_mod  = map(train, \(df) gam(logratio ~ s(range, k = 30), sp = 10e-6, data = df))) |> 
  mutate(
    rmse_linear = map2_dbl(linear_mod, test, \(mod, df) rmse(model = mod, data = df)),
    rmse_smooth = map2_dbl(smooth_mod, test, \(mod, df) rmse(model = mod, data = df)),
    rmse_wiggly = map2_dbl(wiggly_mod, test, \(mod, df) rmse(model = mod, data = df)))
```

I’m mostly focused on RMSE as a way to compare these models, and the plot below shows the distribution of RMSE values for each candidate model.

```{r}
cv_df |> 
  select(starts_with("rmse")) |> 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") |> 
  mutate(model = fct_inorder(model)) |> 
  ggplot(aes(x = model, y = rmse)) + geom_violin()
```


## Example: Child Growth

import the data

```{r}
child_growth = read_csv("./data/nepalese_children.csv")
```

weight vs arm circumference
```{r}
child_growth |> 
  ggplot(aes(x = weight, y = armc)) + 
  geom_point(alpha = .5)
```

```{r}
child_growth =
  child_growth |> 
  mutate(weight_cp7 = (weight > 7) * (weight - 7))
```

Fit the models I care about

```{r}
linear_mod = lm(armc ~ weight, data = child_growth)
pwl_mod    = lm(armc ~ weight + weight_cp7, data = child_growth)
smooth_mod = gam(armc ~ s(weight), data = child_growth)
```

As before, I’ll plot the three models to get intuition for goodness of fit.

```{r}
child_growth |> 
  gather_predictions(linear_mod, pwl_mod, smooth_mod) |> 
  mutate(model = fct_inorder(model)) |> 
  ggplot(aes(x = weight, y = armc)) + 
  geom_point(alpha = .5) +
  geom_line(aes(y = pred), color = "red") + 
  facet_grid(~model)
```

It’s not clear which is best – the linear model is maybe too simple, but the piecewise and non-linear models are pretty similar! Better check prediction errors using the same process as before – again, since I want to fit a gam model, I have to convert the resample objects produced by crossv_mc to dataframes, but wouldn’t have to do this if I only wanted to compare the linear and piecewise models.

```{r}
cv_df =
  crossv_mc(child_growth, 100) |> 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble))
```

Next I’ll use mutate + map & map2 to fit models to training data and obtain corresponding RMSEs for the testing data.

```{r}
cv_df = 
  cv_df |> 
  mutate(
    linear_mod  = map(train, \(df) lm(armc ~ weight, data = df)),
    pwl_mod     = map(train, \(df) lm(armc ~ weight + weight_cp7, data = df)),
    smooth_mod  = map(train, \(df) gam(armc ~ s(weight), data = as_tibble(df)))) |> 
  mutate(
    rmse_linear = map2_dbl(linear_mod, test, \(mod, df) rmse(model = mod, data = df)),
    rmse_pwl    = map2_dbl(pwl_mod, test, \(mod, df) rmse(model = mod, data = df)),
    rmse_smooth = map2_dbl(smooth_mod, test, \(mod, df) rmse(model = mod, data = df)))
```

Violin plot of RMSEs

```{r}
cv_df |> 
  select(starts_with("rmse")) |> 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") |> 
  mutate(model = fct_inorder(model)) |> 
  ggplot(aes(x = model, y = rmse)) + geom_violin()
```






