---
title: "Linear models"
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)

library(rvest)
library(p8105.datasets)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.color = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_color_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d

set.seed(1)
```


## Import data

```{r}
data("nyc_airbnb")

nyc_airbnb = 
  nyc_airbnb |> 
  mutate(stars = review_scores_location / 2) |> 
  rename(
    borough = neighbourhood_group,
    neighborhood = neighbourhood) |> 
  filter(borough != "Staten Island") |> 
  select(price, stars, borough, neighborhood, room_type)
```

## Fit a model

```{r}
fit = lm(price ~ stars + borough, data = nyc_airbnb)
```


Let's look at the result...

```{r, include = FALSE}
fit

summary(fit)
summary(fit)$coef
coef(fit)
residuals(fit)
```

Let's look at the result better...

```{r}
broom::glance(fit)

broom::tidy(fit)
```

and better...

```{r}
fit |> 
  broom::tidy() |> 
  select(term, estimate, p.value) |> 
  mutate(term = str_replace(term, "^borough", "Borough: ")) |> 
  knitr::kable(digits = 3)
```

## Be in control of factors

```{r}
nyc_airbnb = 
  nyc_airbnb |> 
  mutate(
    borough = fct_infreq(borough),
    room_type = fct_infreq(room_type))

fit = lm(price ~ stars + borough, data = nyc_airbnb)
```

Look at the result again

```{r}
broom::tidy(fit)
broom::glance(fit)
```

## Diagnostics

```{r}
modelr::add_residuals(nyc_airbnb, fit)
modelr::add_predictions(nyc_airbnb, fit)
```

```{r}
nyc_airbnb |> 
  modelr::add_residuals(fit) |> 
  ggplot(aes(x = borough, y = resid)) + geom_violin()
```

```{r}
nyc_airbnb |> 
  modelr::add_residuals(fit) |> 
  ggplot(aes(x = stars, y = resid)) + geom_point()
```

## Hypothesis tests

This does t-test by default

```{r}
fit %>% 
  broom::tidy()
```

what about the significant of `borough`.

```{r}
fit_null = lm(price ~ stars + borough, data = nyc_airbnb)
fit_alt = lm(price ~ stars + borough + room_type, data = nyc_airbnb)

anova(fit_null, fit_alt) |> 
  broom::tidy()
```


## Nest data, fit models

This is pretty formal and also complex

```{r}
nyc_airbnb |> 
  lm(price ~ stars * borough + room_type * borough, data = _) |> 
  broom::tidy() |> 
  knitr::kable(digits = 3)
```

This is more exploratory but maybe easier to understand

```{r}
nest_lm_res =
  nyc_airbnb |>
  nest(data = -borough) |>  # 把每个 borough 的数据单独打包成一个小表
  mutate(
    models = map(data, \(df) lm(price ~ stars + room_type, data = df)),  # 在每个小表上建一个模型
    results = map(models, broom::tidy)
  ) |>
  unnest(results)
```

Let's see the results

```{r}
nest_lm_res |> 
  select(borough, term, estimate) |> 
  mutate(term = fct_inorder(term)) |> 
  pivot_wider(
    names_from = term, values_from = estimate) |> 
  knitr::kable(digits = 3)
```

Let's nest even more ...

```{r}
manhattan_airbnb =
  nyc_airbnb |> 
  filter(borough == "Manhattan")

manhattan_nest_lm_res =
  manhattan_airbnb |> 
  nest(data = -neighborhood) |> 
  mutate(
    models = map(data, \(df) lm(price ~ stars + room_type, data = df)),
    results = map(models, broom::tidy)) |> 
  select(-data, -models) |> 
  unnest(results)
```

```{r}
manhattan_nest_lm_res |> 
  filter(str_detect(term, "room_type")) |> 
  ggplot(aes(x = neighborhood, y = estimate)) + 
  geom_point() + 
  facet_wrap(~term) + 
  theme(axis.text.x = element_text(angle = 80, hjust = 1))
```

```{r}
manhattan_airbnb |> 
  lme4::lmer(price ~ stars + room_type + (1 + room_type | neighborhood), data = _) |> 
  broom.mixed::tidy()
```


## Binary outcomes

```{r}
baltimore_df = 
  read_csv("data/homicide-data.csv") |> 
  filter(city == "Baltimore") |> 
  mutate(
    resolved = as.numeric(disposition == "Closed by arrest"),
    victim_age = as.numeric(victim_age),
    victim_race = fct_relevel(victim_race, "White")) |> 
  select(resolved, victim_age, victim_race, victim_sex)
```

```{r}
fit_logistic = 
  baltimore_df |> 
  glm(resolved ~ victim_age + victim_race + victim_sex, data = _, family = binomial()) 
```

```{r}
fit_logistic |> 
  broom::tidy() |> 
  mutate(OR = exp(estimate)) |>
  select(term, log_OR = estimate, OR, p.value) |> 
  knitr::kable(digits = 3)
```

```{r}
baltimore_df |> 
  modelr::add_predictions(fit_logistic) |> 
  mutate(fitted_prob = boot::inv.logit(pred))
```



